\LARGE \textbf{Peer-reviewed international conferences}

\normalsize
\textbf{Bozhko, Y.}, and Heinrich, E. (2011). Concept Map-Based Framework for
Learner-Centered Knowledge Management in ePortfolios. In Proceedings of
The 11th IEEE International Conference on Advanced Learning Technologies 2011.
Athens, GA, USA.

\textbf{Bozhko, Y.}, and Heinrich, E. (2011). Enhancing ePortfolio Systems to
Better Support Lifelong Learning in Universities: Students' Perspective. In
Proceedings of World Conference on Educational Multimedia, Hypermedia and
Telecommunications 2011 (pp. 1912-1917). Chesapeake, VA: AACE.

\textbf{Bozhko, Y.}, and Heinrich, E. (2011). Academic Perspective on Enhancing
ePortfolio Systems to Better Support Lifelong Learning in Universities. In S.
Barton (Ed.), Proceedings of Global Learn Asia Pacific 2011 (pp. 137-142).
Melbourne, Australia: AACE.

﻿\textbf{Bozhko, Y.}, and Heinrich, E. (2010). Towards a Lifelong Learning
Environment in Universities. In Z. Abas et al. (Eds.), Proceedings of Global
Learn Asia Pacific 2010 (pp. 2038-2043). Penang, Malaysia: AACE. 

\textbf{Bozhko, Y.} (2010). Towards an institutional lifelong learning
environment. DEANZ Conference 2010. Wellington, New Zealand.

\textbf{Bozhko, Y.} (2009). Lifelong learning supported by ePortfolio processes.
7th International ePortfolio Conference. London, UK.

\LARGE \textbf{Other publications}

\normalsize
﻿Heinrich, E., and \textbf{Bozhko, Y.} (2011). The Role of Institutions in
Creating Student-Focused Virtual Learning Spaces with ePortfolio Systems. In M. Keppell,
K. Souter, and M. Riddle (Eds.), Physical and Virtual Learning Spaces in Higher
Education: Concepts for the Modern Learning Environment. IGI Global.

\textbf{Bozhko, Y.} (2011). Concept Maps for Learner-Centered Knowledge
Management in ePortfolios. 9th New Zealand Computer Science Research Student Conference
(NZCSRSC) 2011. Palmerston North, New Zealand.

\textbf{Bozhko, Y.} (2010). Towards an Institutional Lifelong Learning
Environment. 8th New Zealand Computer Science Research Student Conference (NZCSRSC) 2010.
Wellington, New Zealand.




