It would be difficult to imagine writing this thesis without help of many
people.

First of all, I would like to thank my main supervisor, Eva Heinrich, who did
an amazing job of keeping me on right track in my research journey; her
experience, guidance, feedback, advice, and criticism were an invaluable source
of inspiration for me.

I would also like to thank my co-supervisors, Mark Brown and David Parsons, who
have provided a valuable educational and methodological perspective on my
research. Their insight and advice were very much appreciated.

Thank you to my parents, who probably still think that my intention to finish
Ph.D. was just another ambitious dream, but nonetheless, have never refused to
support me in all my beginning. I am really grateful for your patience.

I am also thankful for the financial support in form of Doctoral Completion
Bursary that I received from Massey University Graduate Research School. It
motivated me to finish my doctoral studies in timely manner.

Thank you to all people and postgraduate students at SEAT, especially Jevon
Wright, Stephen Lean, Michele Wagner, and Dilantha Punchihewa, for all kinds of
support and creating a great work environment. Also, thank you to Erin Vennell
for taking a difficult responsibility of proof reading my thesis.

On a personal note, I would like to thank Tony Paterson, who planted a seed of
thought in my head that PhD was a dream worthwhile to pursue. Without his
support I would have never attempted to go so far from home.

Finally, I am thankful to Graham Jenson, who I owe my sanity during these years
of Ph.D. study. Without his love, encouragement, and care this Ph.D. would have
been a much less enjoyable experience.
